### 2024-03-27 content/chapter_5.tex
:pencil: Update chapter 4

\chapter{Channel Shaping and Interference Alignment using \glsfmtshort{ris}: From Diagonal to Beyond}\label{ch:channel_shaping}

@@

	The answer indeed depends on the hardware architecture and scattering model.
	In conventional (a.k.a. diagonal) \gls{ris}, each scattering element is tuned by a dedicated impedance and acts as an \emph{individual} phase shifter \cite{Wu2020}.
	The concept is generalized to \gls{bd}-\gls{ris} \cite{Shen2020a,Li2023b} which groups adjacent elements using passive components.
	This allows \emph{cooperative} scattering --- wave impinging on one element can propagate within the circuit and depart partially from any element in the same group.
	\gls{bd}-\gls{ris} can thus control both amplitude and phase of the reflected wave, generalizing the scattering matrix from diagonal with unit-magnitude entries to block diagonal with  unitary blocks.
	Its benefit has been recently shown in receive power maximization \cite{Nerini2023,Santamaria2023,Fang2023,Nerini2023a}, transmit power minimization \cite{Zhou2023}, and rate maximization \cite{Zhou2023,Nerini2023a,Li2023d,Bartoli2023,Li2023c}.
	Practical issues such as channel estimation \cite{Li2023e} and mutual coupling \cite{Li2023f} have also been investigated.
	Therefore, \gls{bd}-\gls{ris} is envisioned as the next generation channel shaper with stronger signal processing flexibility \cite{Li2023g}.

	% Attempts to characterize the channel shaping capability can be classified into \emph{singular value centric} and \emph{power centric} methods.
	% Channel shaping is different from passive beamforming as it seeks to modify the inherent properties of the channel itself, allowing one to decouple \gls{ris} and transceiver design.
	Channel shaping is different from passive beamforming as it seeks to modify the inherent properties of the channel itself.
	This allows one to decouple the \gls{ris}-transceiver design and explore the fundamental limits of channel manipulation.
	For example, diagonal \gls{ris} has been proved useful for improving channel power \cite{Ning2020}, degree of freedom \cite{Ozdogan2020a,Li2023h}, condition number \cite{Zheng2022,Huang2023}, and effective rank \cite{ElMossallamy2021,Meng2023} in \gls{mimo}.
	In contrast, \gls{bd}-\gls{ris} can provide a higher channel power but existing results are limited to \gls{siso}\footnote{In terms of channel shaping, single-stream \gls{mimo} with given precoder and combiner \cite{Nerini2023} is equivalent to \gls{siso}.}. \cite{Nerini2023} and \gls{miso} \cite{Santamaria2023}.
	While these studies offer promising glimpses into the channel shaping potential, a comprehensive understanding of the capabilities and limitations is desired, and a universal design framework is missing.
	This paper aims to answer the channel shaping question through theoretical analysis and numerical optimization.
	The contributions are summarized below.

	First, we quantify the capability of a \gls{bd}-\gls{ris} to reshape the \gls{mimo} point-to-point channel in terms of singular values.
	The \emph{Pareto frontiers} are characterized by optimizing the {weighted sum of singular values}, where the weights can be positive, zero, or negative.
	% This generalizes most singular value metrics and provides a powerful design framework.
	The resulting singular value region generalizes most relevant metrics and provides an intuitive channel shaping benchmark.
	We then discuss some analytical singular value bounds in rank-deficient and fully-connected scenarios, which help to demystify the gain from off-diagonal entries.
	This is the first paper to answer the channel shaping question and highlight the \gls{bd}-\gls{ris} gain from a Pareto perspective.

	Second, we propose a \gls{rcg} algorithm adapted from \cite{Abrudan2008,Abrudan2009} for smooth optimization problems of asymmetric \gls{bd}-\gls{ris} with arbitrary group size.
	Specifically, block-wise update is performed along the geodesics\footnote{A geodesic refers to the shortest path between two points in a Riemannian manifold.} of the Stiefel manifold, which are expressed compactly by the exponential map \cite{Edelman1998}.
	It features lower complexity and faster convergence than general manifold optimization \cite{Absil2009,Pan2022d}, and solves the Pareto singular value problem.
	This is the first paper to tailor an efficient optimization framework for asymmetric \gls{bd}-\gls{ris}.

	Third, we tackle \gls{bd}-\gls{ris} \gls{mimo} rate maximization with two solutions: a local-optimal approach through \gls{ao} and a low-complexity approach over channel shaping.
	The former updates active and passive beamformers by eigenmode transmission and \gls{rcg} algorithm, respectively.
	The latter suboptimally decouples both blocks, recasts the shaping problem as channel power maximization, and solves it iteratively in closed form.
	Interestingly, the gap in between vanishes as \gls{bd}-\gls{ris} evolves from diagonal (single-connected) to unitary (fully-connected).
	It suggests channel shaping offers a promising low-complexity solution for joint \gls{ris}-transceiver designs.

	% Third, we propose a closed-form iterative algorithm for power centric channel shaping problems.
	% The idea is to successively approximate the quadratic objective by a sequence of affines and solve the local problems by \gls{svd}.
	% Case studies are conducted for channel power maximization in \gls{pc} and leakage interference minimization in \gls{ic}.

	Fourth, extensive simulations reveal that the performance gain from \gls{bd}-\gls{ris} increases with group size and \gls{mimo} dimensions.
	In terms of channel power, fully-connected \gls{bd}-\gls{ris} boosts up to 62\% and 270\% over diagonal \gls{ris} in $1 \times 1$ and $4 \times 4$ \gls{mimo} under independent Rayleigh fading, respectively.
	The superiority stems from stronger \emph{subchannel rearrangement} and \emph{subspace alignment} capabilities empowered by in-group cooperation.
	It emphasizes the importance of using \gls{bd}-\gls{ris} in large-scale \gls{mimo} systems.

@@

	% \begin{figure}[!t]
	% 	\centering
	% 	\subfloat[\gls{ris} Elements, $N^\mathrm{T} = 8$\label{fg:pc_rate_sx}]{
	% 		\resizebox{0.48\columnwidth}{!}{
	% 			\input{assets/chapter_5/pc_rate_sx.tex}
	% 		}
	% 	}
	% 	\subfloat[Transmit Antenna, $N^\mathrm{S} = 256$\label{fg:pc_rate_tx}]{
	% 		\resizebox{0.48\columnwidth}{!}{
	% 			\input{assets/chapter_5/pc_rate_tx.tex}
	% 		}
	% 	}
	% 	\caption{Average achievable rate versus group size $L$. $N^\mathrm{R} = 4$, $(\Lambda^\mathrm{D}, \Lambda^\mathrm{F}, \Lambda^\mathrm{B}) = (65, 54, 46) \unit{dB}$.}
	% 	\label{fg:pc_rate}
	% \end{figure}

	% Fig.~\ref{fg:pc_rate_sx} illustrates how \gls{ris} configuration influences the \gls{mimo} \gls{pc} achievable rate.
	% To ensure a \qty{20}{bit/s/Hz} transmission, an \gls{snr} of \qty{13.5}{\dB} is required for a 8T4R system.
	% This value decreases to \qty{12.5}{\dB} (resp. \qty{8}{\dB}) when 32- (resp. 256-) element diagonal \gls{ris} is present.
	% If tetrads can be formed in \gls{bd}-\gls{ris}, the \gls{snr} can be reduced by another \qty{20}{\percent} (resp. \qty{44}{\percent}).
	% Further increase in $L$ yields a marginal gain and incurs $\mathcal{O}(L^2)$ connections.
	% We thus conclude dyadic or tetradic \gls{bd}-\gls{ris} usually strike a good balance between performance and complexity.

	% \begin{figure}[!t]
	% 	\centering
	% 	\resizebox{0.65\columnwidth}{!}{
	% 		\input{assets/chapter_5/power_sx.tex}
	% 	}
	% 	\caption{Average channel power versus \gls{ris} elements $N^\mathrm{S}$ and group size $L$. $(N^\mathrm{T}, N^\mathrm{R}) = (8, 4)$, $(\Lambda^\mathrm{D}, \Lambda^\mathrm{F}, \Lambda^\mathrm{B}) = (65, 54, 46) \unit{dB}$.}
	% 	\label{fg:power_sx}
	% \end{figure}

	% \begin{figure}[!t]
	% 	\centering
	% 	\subfloat[Without Direct Link\label{fg:pc_power_bond_nd}]{
	% 		\resizebox{0.48\columnwidth}{!}{
	% 			\input{assets/chapter_5/pc_power_bond_nd.tex}
	% 		}
	% 	}
	% 	\subfloat[With Direct Link\label{fg:pc_power_bond_hd}]{
	% 		\resizebox{0.48\columnwidth}{!}{
	% 			\input{assets/chapter_5/pc_power_bond_hd.tex}
	% 		}
	% 	}
	% 	\caption{Average channel power versus \gls{ris} group size $L$. $(N^\mathrm{T}, N^\mathrm{S}, N^\mathrm{R}) = (8, 256, 4)$, $(\Lambda^\mathrm{D}, \Lambda^\mathrm{F}, \Lambda^\mathrm{B}) = (65, 54, 46) \unit{dB}$.}
	% 	\label{fg:pc_power_bond}
	% \end{figure}
	% % Fig.~\ref{fg:power_sx} shows that, apart from numerous reflecting elements, a sufficiently large group size can help
	% % Fig.~\ref{fg:power_sx} shows that increasing the group size can improve the channel power especially for a large \gls{ris}.
	% Fig.~\ref{fg:power_sx} shows that, apart from adding reflecting elements $N^\mathrm{S}$, increasing the group size $L$ also improves the channel power.
	% % The behavior is more obvious for a large \gls{ris}.
	% This behavior is more pronounced for a large \gls{ris}.
	% For example, the gain of pairwise connection is \qty{2.8}{\percent} for $N^\mathrm{S} = 16$ and \qty{28}{\percent} for $N^\mathrm{S} = 256$.
	% % \gls{bd}-\gls{ris} provides a higher channel power than conventional \gls{ris}.
	% % As $L$ increases from 1 to 2,
	% It implies that the channel shaping capability of \gls{bd}-\gls{ris} scales with group size $L$.

@@

>>>>

\chapter{Channel Shaping using \glsfmtshort{ris}: From Diagonal Model to Beyond}\label{ch:channel_shaping}

@@

	The answer depends heavily on the hardware architecture and scattering model.
	In diagonal \gls{ris}, each scattering element is tuned by a dedicated impedance and acts as an \emph{individual} phase shifter \cite{Wu2020}.
	The concept is generalized to \gls{bd}-\gls{ris} \cite{Shen2020a,Li2023b} which physically connects adjacent elements using passive components.
	This allows \emph{cooperative} scattering --- wave impinging on one element can propagate within the circuit and depart partially from any element in the same group.
	\gls{bd}-\gls{ris} can thus control both amplitude and phase of the reflected wave, generalizing the scattering matrix from diagonal with unit-magnitude entries \eqref{eq:diagonal_scattering_matrix} to block diagonal with unitary blocks \eqref{eq:bd_scattering_matrix}.
	Its benefit has been recently shown in receive power maximization \cite{Nerini2023,Santamaria2023,Fang2023,Nerini2023a}, transmit power minimization \cite{Zhou2023}, and rate maximization \cite{Zhou2023,Nerini2023a,Li2023d,Bartoli2023,Li2023c}.
	Practical issues such as channel estimation \cite{Li2023e} and mutual coupling \cite{Li2023f} have also been investigated.
	Therefore, \gls{bd}-\gls{ris} is envisioned as the next generation channel shaper with stronger signal processing flexibility \cite{Li2023g}.

	% Attempts to characterize the channel shaping capability can be classified into \emph{singular value centric} and \emph{power centric} methods.
	% Channel shaping is different from passive beamforming as it seeks to modify the inherent properties of the channel itself, allowing one to decouple \gls{ris} and transceiver design.
	Channel shaping is different from passive beamforming discussed in Chapters \ref{ch:ris_aided_swipt} and \ref{ch:riscatter} because \gls{ris} is exploited as a stand-alone device to modify the inherent properties of the wireless channel.
	This allows one to decouple the \gls{ris}-transceiver design and explore the fundamental limits of wave manipulation.
	Relevant design criteria can be roughly classified into two categories:
	\begin{itemize}
		\item \emph{Singular value centric:} Focus on the singular values of the equivalent channel matrix, which are closely related to the channel capacity and the diversity-multiplexing trade-off. Relevant designs are more complicated since they involve the entire channel matrix.
		\item \emph{Power centric:} Focus on the second-order characteristics of the channel matrix, which are useful in \gls{miso} but less informative in \gls{mimo}. Relevant designs are more practical and straightforward.
	\end{itemize}
	The former has been studied in terms of minimum singular value \cite{ElMossallamy2021}, effective rank \cite{ElMossallamy2021,Meng2023}, condition number \cite{Zheng2022,Huang2023} in \gls{pc}, and degree of freedom \cite{Bafghi2022,Zheng2023,Chae2023} in \gls{ic}.
	The latter has been studied in terms of channel power \cite{Wu2019,Nerini2023,Santamaria2023} in \gls{pc} and leakage interference \cite{Santamaria2023a} in \gls{ic}.
	Although the results are insightful, the big picture is far from complete since there exists neither a systematic channel shaping study nor a universal \gls{ris} optimization framework.
	To the best of our knowledge, the most general channel shaping scenario (\gls{bd}-\gls{ris}-aided \gls{mimo}) has not been studied, and the great potential of \gls{bd}-\gls{ris} has not been fully exploited or understood.
	This chapter aims for a comprehensive answer to the channel shaping question through theoretical analysis and numerical optimization.
	The contributions are summarized below.

	First, we exploit the Riemannian geometry of the Stiefel manifold to design a novel geodesic\footnote{A geodesic refers to the shortest path between two points in a Riemannian manifold.} \gls{rcg} algorithm for \gls{bd}-\gls{ris} optimization.
	This method modified from \cite{Abrudan2008,Abrudan2009} not only features lower complexity and faster convergence than general non-geodesic approach \cite{Absil2009,Pan2022d}, but also fits arbitrary group size and any smooth \gls{ris} optimization problem.
	Specifically, block-wise rotational update is performed along the geodesics of the Stiefel manifold, which are expressed compactly by the exponential map \cite{Edelman1998}.
	This is the first work to tailor an efficient optimization framework for asymmetric \gls{bd}-\gls{ris}.

	Second, we quantify the capability of a \gls{bd}-\gls{ris} to redistribute the singular values of \gls{mimo}-\gls{pc}.
	The \emph{Pareto frontiers} are characterized by optimizing the {weighted sum of singular values}, where the weights can be positive, zero, or negative.
	This problem is solved by the proposed geodesic \gls{rcg} algorithm.
	The resulting achievable singular value region generalizes most relevant metrics and provides an intuitive channel shaping benchmark.
	We also demystify the \gls{bd}-\gls{ris} gain from off-diagonal entries and derive some analytical singular value bounds for rank-deficient \gls{mimo} and fully-connected (a.k.a. unitary) \gls{ris}.
	This is the first work to comprehensively answer the channel shaping question and fully unveil the potential or \gls{bd}-\gls{ris}.

	Third, we tackle \gls{bd}-\gls{ris}-aided \gls{mimo}-\gls{pc} rate maximization with two solutions: a local-optimal approach through \gls{ao} and a low-complexity approach over channel shaping.
	The former updates active and passive beamforming by eigenmode transmission and \gls{rcg} algorithm, respectively.
	The latter suboptimally decouples both blocks, recasts the shaping problem as channel power maximization, and solves it iteratively in closed form.
	Interestingly, the gap in between vanishes as \gls{bd}-\gls{ris} evolves from diagonal to unitary.
	It suggests channel shaping decouples the \gls{ris}-transceiver design and offers a promising low-complexity solution.
	This is the first work to discuss the implications of channel shaping in joint active-passive beamforming optimization.

	Fourth, we extend the aforementioned approaches to \gls{mimo}-\gls{ic} where the \gls{bd}-\gls{ris} is used for leakage interference minimization and \gls{wsr} maximization.
	In the former case, we update the receive combiner, transmit precoder, and scattering matrix iteratively in closed forms with optimality proof.
	In the latter case, we alternatively updates the transmit precoder by bisection and the scattering matrix by \gls{rcg} algorithm.
	We also compare the impact of \gls{ris} on \gls{pc} and \gls{ic}.
	% provide a comprehensive comparison between the two approaches and discuss the implications of channel shaping in \gls{ic}.

	% Fifth, extensive simulations reveal that the performance gain from \gls{bd}-\gls{ris} increases with group size and \gls{mimo} dimensions.
	% In terms of channel power, fully-connected \gls{bd}-\gls{ris} boosts up to 62\% and 270\% over diagonal \gls{ris} in $1 \times 1$ and $4 \times 4$ \gls{mimo} under independent Rayleigh fading, respectively.
	% The superiority stems from stronger \emph{subchannel rearrangement} and \emph{subspace alignment} capabilities empowered by in-group cooperation.
	% It emphasizes the importance of using \gls{bd}-\gls{ris} in large-scale \gls{mimo} systems.

@@

	% \begin{figure}[H]
	% 	\centering
	% 	\subfloat[\gls{ris} Elements, $N^\mathrm{T} = 8$\label{fg:pc_rate_sx}]{
	% 		\resizebox{0.48\columnwidth}{!}{
	% 			\input{assets/chapter_5/pc_rate_sx.tex}
	% 		}
	% 	}
	% 	\subfloat[Transmit Antenna, $N^\mathrm{S} = 256$\label{fg:pc_rate_tx}]{
	% 		\resizebox{0.48\columnwidth}{!}{
	% 			\input{assets/chapter_5/pc_rate_tx.tex}
	% 		}
	% 	}
	% 	\caption{Average achievable rate versus group size $L$. $N^\mathrm{R} = 4$, $(\Lambda^\mathrm{D}, \Lambda^\mathrm{F}, \Lambda^\mathrm{B}) = (65, 54, 46) \unit{dB}$.}
	% 	\label{fg:pc_rate}
	% \end{figure}

	% Fig.~\ref{fg:pc_rate_sx} illustrates how \gls{ris} configuration influences the \gls{mimo}-\gls{pc} achievable rate.
	% To ensure a \qty{20}{bit/s/Hz} transmission, an \gls{snr} of \qty{13.5}{\dB} is required for a 8T4R system.
	% This value decreases to \qty{12.5}{\dB} (resp. \qty{8}{\dB}) when 32- (resp. 256-) element diagonal \gls{ris} is present.
	% If tetrads can be formed in \gls{bd}-\gls{ris}, the \gls{snr} can be reduced by another \qty{20}{\percent} (resp. \qty{44}{\percent}).
	% Further increase in $L$ yields a marginal gain and incurs $\mathcal{O}(L^2)$ connections.
	% We thus conclude dyadic or tetradic \gls{bd}-\gls{ris} usually strike a good balance between performance and complexity.

	% \begin{figure}[H]
	% 	\centering
	% 	\resizebox{0.65\columnwidth}{!}{
	% 		\input{assets/chapter_5/power_sx.tex}
	% 	}
	% 	\caption{Average channel power versus \gls{ris} elements $N^\mathrm{S}$ and group size $L$. $(N^\mathrm{T}, N^\mathrm{R}) = (8, 4)$, $(\Lambda^\mathrm{D}, \Lambda^\mathrm{F}, \Lambda^\mathrm{B}) = (65, 54, 46) \unit{dB}$.}
	% 	\label{fg:power_sx}
	% \end{figure}

	% \begin{figure}[H]
	% 	\centering
	% 	\subfloat[Without Direct Link\label{fg:pc_power_bond_nd}]{
	% 		\resizebox{0.48\columnwidth}{!}{
	% 			\input{assets/chapter_5/pc_power_bond_nd.tex}
	% 		}
	% 	}
	% 	\subfloat[With Direct Link\label{fg:pc_power_bond_hd}]{
	% 		\resizebox{0.48\columnwidth}{!}{
	% 			\input{assets/chapter_5/pc_power_bond_hd.tex}
	% 		}
	% 	}
	% 	\caption{Average channel power versus \gls{ris} group size $L$. $(N^\mathrm{T}, N^\mathrm{S}, N^\mathrm{R}) = (8, 256, 4)$, $(\Lambda^\mathrm{D}, \Lambda^\mathrm{F}, \Lambda^\mathrm{B}) = (65, 54, 46) \unit{dB}$.}
	% 	\label{fg:pc_power_bond}
	% \end{figure}
	% % Fig.~\ref{fg:power_sx} shows that, apart from numerous reflecting elements, a sufficiently large group size can help
	% % Fig.~\ref{fg:power_sx} shows that increasing the group size can improve the channel power especially for a large \gls{ris}.
	% Fig.~\ref{fg:power_sx} shows that, apart from adding reflecting elements $N^\mathrm{S}$, increasing the group size $L$ also improves the channel power.
	% % The behavior is more obvious for a large \gls{ris}.
	% This behavior is more pronounced for a large \gls{ris}.
	% For example, the gain of pairwise connection is \qty{2.8}{\percent} for $N^\mathrm{S} = 16$ and \qty{28}{\percent} for $N^\mathrm{S} = 256$.
	% % \gls{bd}-\gls{ris} provides a higher channel power than conventional \gls{ris}.
	% % As $L$ increases from 1 to 2,
	% It implies that the channel shaping capability of \gls{bd}-\gls{ris} scales with group size $L$.

@@

\begin{section}{\glsfmtshort{mimo}-\glsfmtshort{ic}}
	\begin{subsection}{Leakage Interference Minimization}
		\begin{mini!}
			{\scriptstyle{\mathbf{\Theta}, \{\mathbf{G}_k\}, \{\mathbf{W}_k\}}}{\mathop{\sum\sum}_{j \neq k} \left\lVert \mathbf{G}_k (\mathbf{H}_{kj}^\mathrm{D} + \mathbf{H}_k^\mathrm{B} \mathbf{\Theta} \mathbf{H}_j^\mathrm{F}) \mathbf{W}_j \right\rVert _{\mathrm{F}}^2}{}{}
			\addConstraint{\mathbf{\Theta}_g^\mathsf{H} \mathbf{\Theta}_g=\mathbf{I}, \quad \forall g}{}{}
			\addConstraint{\mathbf{G}_k \mathbf{G}_k^\mathsf{H}=\mathbf{I}, \quad \mathbf{W}_k^\mathsf{H} \mathbf{W}_k=\mathbf{I}, \quad \forall k.}{}{}
		\end{mini!}
		The non-convex problem can be solved by \gls{bcd} method.
		For a given $\mathbf{\Theta}$, it reduces to conventional linear beamforming problem, for which an iterative algorithm alternating between the original and reciprocal networks is proposed in \cite{Gomadam2011,Clerckx2013}.
		At iteration $r$, the combiner at receiver $k$ is updated as
		\begin{equation}
			\mathbf{G}_k^{(r)} = {\mathbf{U}_{k,N}^{(r-1)}}^\mathsf{H},
		\end{equation}
		where $\mathbf{U}_{k,N}^{(r-1)}$ is the eigenvectors corresponding to $N$ smallest eigenvalues of interference covariance matrix $\mathbf{Q}_k^{(r-1)} = \sum_{j \ne k} \mathbf{H}_{kj} \mathbf{W}_j^{(r-1)} {\mathbf{W}_j^{(r-1)}}^\mathsf{H} \mathbf{H}_{kj}^\mathsf{H}$.
		The precoder at transmitter $j$ is updated as
		\begin{equation}
			\mathbf{W}_j^{(r)} = \bar{\mathbf{U}}_{j,N}^{(r)},
		\end{equation}
		where $\bar{\mathbf{U}}_{j,N}^{(r)}$ corresponds to interference covariance matrix $\bar{\mathbf{Q}}_j^{(r)} = \sum_{k \ne j} \mathbf{H}_{kj}^\mathsf{H} {\mathbf{G}_k^{(r)}}^\mathsf{H} \mathbf{G}_k^{(r)} \mathbf{H}_{kj}$ in the reciprocal network.
		Once $\{\mathbf{G}_k\}$ and $\{\mathbf{W}_k\}$ are determined, we define $\bar{\mathbf{H}}_{kj}^\mathrm{D} \triangleq \mathbf{G}_k \mathbf{H}_{kj}^\mathrm{D} \mathbf{W}_j$, $\bar{\mathbf{H}}_k^\mathrm{B} \triangleq \mathbf{G}_k \mathbf{H}_k^\mathrm{B}$, and $\bar{\mathbf{H}}_j^\mathrm{F} \triangleq \mathbf{H}_j^\mathrm{F} \mathbf{W}_j$.
		The \gls{bd} \gls{ris} subproblem reduces to
		\begin{mini!}
			{\scriptstyle{\mathbf{\Theta}}}{\mathop{\sum\sum}_{j \neq k} \left\lVert (\bar{\mathbf{H}}_{kj}^\mathrm{D} + \bar{\mathbf{H}}_k^\mathrm{B} \mathbf{\Theta} \bar{\mathbf{H}}_j^\mathrm{F}) \right\rVert _{\mathrm{F}}^2}{\label{op:ic_interference_ris}}{}
			\addConstraint{\mathbf{\Theta}_g^\mathsf{H} \mathbf{\Theta}_g=\mathbf{I}, \quad \forall g.}{}{}
		\end{mini!}

		\begin{proposition}
			Start from any $\mathbf{\Theta}^{(0)}$, the sequence
			\begin{equation}
				\mathbf{\Theta}_g^{(r+1)} = \mathbf{U}_g^{(r)} \mathbf{V}_g^{(r)}, \quad \forall g
			\end{equation}
			converges to a stationary point of \eqref{op:ic_interference_ris}, where $\mathbf{U}_g^{(r)}$ and $\mathbf{V}_g^{(r)}$ are left and right singular matrix of
			\begin{equation}
				\mathbf{M}_g^{(r)} = \mathop{\sum\sum}_{j \neq k} \bigl(\mathbf{B}_{k,g} \mathbf{\Theta}_g^{(r)} \mathbf{H}_{j,g}^\mathrm{F} - {\mathbf{H}_{k,g}^\mathrm{B}}^\mathsf{H} \mathbf{D}_{kj,g}^{(r)}\bigr) {\mathbf{H}_{j,g}^\mathrm{F}}^\mathsf{H},
			\end{equation}
			where $\mathbf{B}_{k,g} = \lambda_1\bigl(\mathbf{H}_{k,g}^\mathrm{B} {\mathbf{H}_{k,g}^\mathrm{B}}^\mathsf{H}\bigr) \mathbf{I} - {\mathbf{H}_{k,g}^\mathrm{B}}^\mathsf{H} \mathbf{H}_{k,g}^\mathrm{B}$ and
			\begin{equation}
				\mathbf{D}_{kj,g}^{(r)} = \mathbf{H}_{jk}^\mathrm{D} + \sum_{g'<g} {\mathbf{H}_{k,g'}^\mathrm{B}}^\mathsf{H} \mathbf{\Theta}_{g'}^{(r+1)} \mathbf{H}_{k,g'}^\mathrm{F} + \sum_{g'>g} {\mathbf{H}_{k,g'}^\mathrm{B}}^\mathsf{H} \mathbf{\Theta}_{g'}^{(r)} \mathbf{H}_{k,g'}^\mathrm{F}.
			\end{equation}
		\end{proposition}
		\begin{proof}
			To be added.
		\end{proof}

		\begin{figure}[!t]
			\centering
			\resizebox{0.65\columnwidth}{!}{
				\input{assets/chapter_5/ic_interference_sx.tex}
			}
			\caption{Average leakage interference versus \gls{ris} elements $N^\mathrm{S}$ and group size $L$. Transmitters and receivers are randomly generated in a disk of radius 50 m centered at the \gls{ris}. $(N^\mathrm{T}, N^\mathrm{R}, N^\mathrm{E}, K) = (8, 4, 3, 5)$, $(\gamma^\mathrm{D}, \gamma^\mathrm{F}, \gamma^\mathrm{B}) = (3, 2.4, 2.4)$, and reference pathloss at \qty{1}{\meter} is \qty{-30}{\dB}.}
			\label{sm:ic_interference_sx}
		\end{figure}
		Fig.~\ref{sm:ic_interference_sx} illustrates how \gls{bd} \gls{ris} helps to reduce the leakage interference.
		In this case, a fully-connected $2^n$-element \gls{bd} \gls{ris} is almost as good as a diagonal $2^{n+2}$-element \gls{ris} in terms of leakage interference.
		Interestingly, the result suggests that \gls{bd} \gls{ris} can achieve a higher \gls{dof} than diagonal \gls{ris} in \gls{mimo}-\gls{ic}, which is not the case in \gls{mimo}-\gls{pc} (as discussed in \ref{sc:pc_rank_deficient}).
	\end{subsection}

	\begin{subsection}{Weighted Sum-Rate Maximization}
		\begin{maxi!}
			{\scriptstyle{\mathbf{\Theta}, \{\mathbf{W}_k\}}}{J_2 = \sum_k \rho_k \log \det \biggl(\mathbf{I} + \mathbf{W}_k \mathbf{H}_{kj}^\mathsf{H} \mathbf{Q}_k^{-1} \mathbf{H}_{kj} \mathbf{W}_k\biggr)}{\label{op:ic_rate}}{\label{ob:ic_rate}}
			\addConstraint{\mathbf{\Theta}_g^\mathsf{H} \mathbf{\Theta}_g=\mathbf{I}, \quad \forall g}{}{}
			\addConstraint{\lVert \mathbf{W}_k \rVert _\mathrm{F}^2 \le P_k. \quad \forall k}{}{}
		\end{maxi!}
		where $\rho_k$ is the weight of user $k$ and $\mathbf{Q}_k$ is the interference-plus-noise covariance matrix
		\begin{equation}
			\mathbf{Q}_k = \sum_{j \ne k} \mathbf{H}_{kj} \mathbf{W}_j {\mathbf{W}_j}^\mathsf{H} \mathbf{H}_{kj}^\mathsf{H} + \eta \mathbf{I}.
		\end{equation}
		For a given $\mathbf{\Theta}$, \eqref{op:ic_rate} reduces to conventional linear beamforming problem, for which a closed-form iterative solution based on mutual information-\gls{mmse} relationship is proposed in \cite{Negro2010}.
		At iteration $r$, the \gls{mmse} combiner at receiver $k$ is
		\begin{equation}
			\mathbf{G}_k^{(r)} = {\mathbf{W}_k^{(r-1)}}^\mathsf{H} \mathbf{H}_{kk}^\mathsf{H} \bigl(\mathbf{Q}_k^{(r-1)} + \mathbf{H}_{kk} \mathbf{W}_k^{(r-1)} {\mathbf{W}_k^{(r-1)}}^\mathsf{H} \mathbf{H}_{kk}^\mathsf{H}\bigr)^{-1},
		\end{equation}
		the corresponding error matrix is
		\begin{equation}
			\mathbf{E}_k^{(r)} = \bigl(\mathbf{I} + {\mathbf{W}_k^{(r-1)}}^\mathsf{H} \mathbf{H}_{kk}^\mathsf{H} \mathbf{Q}_k^{(r-1)} \mathbf{H}_{kk} \mathbf{W}_k^{(r-1)}\bigr)^{-1},
		\end{equation}
		the \gls{mse} weight is
		\begin{equation}
			\mathbf{\Omega}_k^{(r)} = \rho_k {\mathbf{E}_k^{(r)}}^{-1},
		\end{equation}
		the Lagrange multiplier is
		% \begin{equation}
		% 	\begin{split}
		% 		\lambda_k^{(r)}
		% 		& = \frac{\sum_{j \ne k} \mathrm{tr}(\mathbf{\Omega}_k^{(r)}\mathbf{F}_{kj}^{(r)} {\mathbf{F}_{kj}^{(r)}}^\mathsf{H}) - \sum_{j \ne k} \mathrm{tr}(\mathbf{\Omega}_j^{(r)}\mathbf{F}_{jk}^{(r)} {\mathbf{F}_{jk}^{(r)}}^\mathsf{H})}{P_k} \\
		% 		& \quad + \frac{\mathrm{tr}(\eta \mathbf{\Omega}_k^{(r)} \mathbf{G}_k^{(r)}{\mathbf{G}_k^{(r)}}^\mathsf{H})}{P_k}
		% 	\end{split}
		% \end{equation}
		\begin{equation}
			\lambda_k^{(r)} = \frac{\mathrm{tr}\bigl(\eta \mathbf{\Omega}_k^{(r)} \mathbf{G}_k^{(r)}{\mathbf{G}_k^{(r)}}^\mathsf{H} + \sum_j \mathbf{\Omega}_k^{(r)}\mathbf{T}_{kj}^{(r)} {\mathbf{T}_{kj}^{(r)}}^\mathsf{H} - \mathbf{\Omega}_j^{(r)}\mathbf{T}_{jk}^{(r)} {\mathbf{T}_{jk}^{(r)}}^\mathsf{H} \bigr)}{P_k},
		\end{equation}
		where $\mathbf{T}_{kj}^{(r)} = \mathbf{G}_k^{(r)} \mathbf{H}_{kj} \mathbf{W}_j^{(r)}$.
		The precoder at transmitter $k$ is
		\begin{equation}
			\mathbf{W}_k^{(r)} = \Bigl(\sum_j \mathbf{H}_{jk}^\mathsf{H} {\mathbf{G}_j^{(r)}}^\mathsf{H} \mathbf{\Omega}_k^{(r)} \mathbf{G}_j^{(r)} \mathbf{H}_{jk} + \lambda_k^{(r)} \mathbf{I} \Bigr)^{-1} \mathbf{H}_{kk}^\mathsf{H} {\mathbf{G}_j^{(r)}}^\mathsf{H} \mathbf{\Omega}_k^{(r)}.
		\end{equation}
		Once $\{\mathbf{W}_k\}$ is determined, the complex derivative of \eqref{ob:ic_rate} w.r.t. \gls{ris} block $g$ is
		\begin{equation}
			\begin{split}
				\frac{\partial J_2}{\partial \mathbf{\Theta}_g^*}
				& = \sum_k \rho_k {\mathbf{H}_{k,g}^\mathrm{B}}^\mathsf{H} \mathbf{Q}_k^{-1} \mathbf{H}_{kk} \mathbf{W}_k \mathbf{E}_k \mathbf{W}_k^\mathsf{H} \\
				& \quad \times \bigl({\mathbf{H}_{k,g}^\mathrm{F}}^\mathsf{H} - \mathbf{H}_{kk}^\mathsf{H} \mathbf{Q}_k^{-1} \sum_{j \ne k} \mathbf{H}_{kj} \mathbf{W}_j \mathbf{W}_j^\mathsf{H} {\mathbf{H}_{j,g}^\mathrm{F}}^\mathsf{H}\bigr).
			\end{split}
			\label{eq:ic_rate_gradient_ris}
		\end{equation}
		The \gls{ris} subproblem can be solved by \gls{rcg} Algorithm~\ref{ag:pc_rate_ris} with \eqref{eq:pc_rate_gradient_ris} replaced by \eqref{eq:ic_rate_gradient_ris}.

		\begin{figure}[!t]
			\centering
			\resizebox{0.65\columnwidth}{!}{
				\input{assets/chapter_5/ic_rate_sx.tex}
			}
			\caption{Average weighted sum-rate versus \gls{snr}, \gls{ris} elements $N^\mathrm{S}$ and group size $L$. $(N^\mathrm{T}, N^\mathrm{R}, N^\mathrm{E}, K) = (8, 4, 3, 5)$, $(\Lambda^\mathrm{D}, \Lambda^\mathrm{F}, \Lambda^\mathrm{B}) = (65, 54, 46) \unit{dB}$, $\rho_k = 1, \ \forall k$.}
			\label{sm:ic_rate_sx}
		\end{figure}

		\begin{figure}[!t]
			\centering
			\resizebox{0.65\columnwidth}{!}{
				\input{assets/chapter_5/ic_rate_user.tex}
			}
			\caption{Average weighted sum-rate versus user pairs $K$, \gls{ris} elements $N^\mathrm{S}$ and group size $L$ at $\mathrm{SNR} = \qty{15}{dB}$. $(N^\mathrm{T}, N^\mathrm{R}, N^\mathrm{E}) = (4, 4, 3)$, $\rho_k = 1, \ \forall k$.}
			\label{sm:ic_rate_user}
		\end{figure}

		A new observation from Fig.~\ref{sm:ic_rate_sx} that the interference alignment capability of \gls{bd} \gls{ris} scales much faster with group size than number of elements.\footnote{The results are not very stable and depend heavily on initialization.}
	\end{subsection}
\end{section}
